A Hitchhiker's Guide to Reproducible Research in R

### 2021-08-03 manuscript.Rmd

Scientists increasingly strive to make research data, materials, and analysis code openly available.
Sharing these types of digital research products has the potential to increase scientific efficiency by enabling researchers to learn from each other and reuse materials, and to increase scientific reliability by facilitating the review and replication of published research results.
To some extent these potential benefits are contingent on whether these digital research products are reproducible. 
Reproducibility can be defined as the ability of anyone to obtain identical results from the *same* data with the *same* computer code [see @Peikert2019 for details].
Reproducibility has long been considered integral to empirical research, whose credibility hinges on its objectivity, meaning "in principle it can be tested and understood by anybody." [@popperLogicScientificDiscovery2002, p. 22][^popper]
Unfortunately, despite increasing commitment to open science practices and good will, many projects can not yet be reproduced by other research teams [@obels2020].
This is because making a research project with all code, data, and materials reproducible comes with a variety of challenges, and there is a dire need for reproducibility standards.
We now have the technology to make all digital products related to a research project available so that a research project can be reproduced with minimal effort.

This tutorial is aimed at researchers who intend to ensure the reproducibility of a research project and are willing to make relevant code, data, and materials available, whether publicly or on request.
We pursue two aims.
First, the present paper introduces the reader to a workflow that maximizes the potential reproducibility of a research project.
The end goal of this workflow is a project that can be completely, automatically, and unambiguously reproduced.
This means that any manual steps, such as copy and pasting results, manual modification of data (e.g., manually removing observations or manually recoding variables),
are replaced with reproducible code. 
The result of this workflow will be a portable research project containing all digital research artefacts,
in which all steps required to produce the (published) manuscript from the raw (and ideally, open) data are well-documented
and automatically reproducible with a single command.

This workflow builds on several technical solutions originating in the software engineering community. 
We discuss how these solution address four common threats to reproducibility [see @Peikert2019 for details] and form a coherent workflow.
We further demonstrate how the R package `repro` supports researchers in adopting this workflow.
The second objective is to discuss opportunities of reproducible workflows to improve not only how research outputs are created and recreated, but also how research itself is conducted.
One such opportunity is a strictly reproducible and unambiguous form of preregistration [@NosekRevolution2018] that builds upon a reproducible workflow, the so-called *preregistration as code* (PAC).
PAC involves preregistering the intended analysis code and the bulk of the manuscript as dynamic document, including the Introduction, Methods, and Results sections.
The resulting dynamic document closely resembles the final manuscript, but uses preliminary ("mock") results, based on the analysis of simulated data, as a placeholder for figures, tables, and statistics. 
After the empirical data are analyzed, the Results are updated, and the Discussion is written.

@@

1.	**Multiple inconsistent versions of code, data, or both**; for example, the data set could have changed over time because outliers were removed at a later stage or an item was later recoded; or the analysis code could have been modified during writing of a paper because a bug was removed at some point in time; it may then be unclear, which version of code and data was used to produce some reported set of results.
2.	**Copy-and-paste errors**; for example, often, results are manually copied from a statistical computing language into a text processor; if a given analysis is re-run and results are manually updated in the text processor, this may inadvertently lead to inconsistencies between the reported result and the reproduced result;
3.	**Undocumented or ambiguous order of code execution**; for example, with multiple data and code files, it may be unclear which scripts should be executed in what order;  or, some of the computational steps are documented (e.g., final analysis), but other steps were conducted manually and not documented (e.g., running a scripts/commands by hand; copy-pasting results from one program to another)
4.	**Ambiguous dependencies**; for example, a given analysis may depend on a specific version of a specific software package, or rely on software that might not be available on a different computer or no longer exist at all; or, a different version of the same software may produce different results.

We have developed a workflow that leverages established tools and practices from software engineering to achieve long-term and cross-platform computational reproducibility of scientific data analyses resting on four pillars that address the aforementioned causes of non-reproducibility [@Peikert2019]:

@@

A version controlled dynamic document with dependency tracking may still rely on external software.
Troubleshooting issues specific to a particular programming language or dependent tool typically requires considerable expertise and pose a threat to reproducibility.
One solution to avoid problems caused by missing or inconsistent external software dependencies is to provide not only the analysis script but all dependent software packages and system libraries via **software management**.
One comprehensive approach to software management is containerization.
"packaging the key elements of the computational  environment  needed  to  run  the  desired  software [makes] the software much easier to use, and the results easier to reproduce" [@silverSoftwareSimplified2017, p. 174].

@@

1. Text in natural language (prose; e.g., a scientific paper or presentation)
2. Executable code (e.g., analyses)
3. Metadata (e.g. title, authors, document format)

R Markdown is a type of dynamic document well-suited to the RStudio user interface.
The natural language text in an R Markdown file can be formatted using Markdown (see @xieMarkdownDefinitiveGuide2019 for technical details and @xieMarkdownCookbook2020 for practical guidance).
The code mostly consists of R-code (although other programming languages are supported, like Python, C++, Fortran, etc).
The following example serves to illustrate the Markdown syntax. 
It shows how to create a heading, a word in bold font, and a list of several items in Markdown:

@@

The metadata the `repro::automate()` function relies on can only express dependencies R packages for the Dockerfile and only trivial dependencies (in the form of "file must exist") for the Makefile. 
The first limitation means that users must rely on software that is either already provided by the base Dockerimage "rocker/verse" or the R packages they list in the metadata.
Other software that users might need like other programming languages, not already installed LaTeX packages etc. have to be added manually.
We plan to add support for commonly used ways to install software beyond R packages via the metadata and `repro::automate()` e.g. for system libraries (via `apt` the Ubuntu package manager), LaTeX packages (via `tlmgr` the Tex Live package manager), python packages (via `pip` the python package manager).
The second limitation is about dependencies.
Make can represent complex dependencies like in this example: A depends on B, which in turn depends on C and D.
If B is missing in this example, Make would know how to recreate it from C and D.
These dependencies and especially how these dependencies should be resolved are difficult to represent in the metadata.
Users therefore have to either "flatten" the dependency structure, by simply stating that A depends on B, C and D therefore leaving out important information or express the dependencies directly within the Makefile.

@@

Using Docker may require you to install software on an operating system that may not be familiar to you.
The images supplied by [@boettigerIntroductionRockerDocker2017], for example, are based on the Ubuntu operating system.
The most convenient way to install software on Ubuntu is through its package manager `apt`.
The following snipped, for example, installs Python but works the same for any other software users might want to install.
If added to `.repro/Dockerfile_manual` the Docker image will have Python installed.

@@

The file `.repro/Makefile_Rmds` contains the automatically generated targets from `repro::automate()` for the R Markdown files.
This file should not be altered manually.
If you are not satisfied with the automatically generated target than simply provide an alternative target in the main `Makefile`.
Targets in the main `Makefile` take precedent.

@@

Defining the research questions and planning data analysis before observing the research outcomes is called preregistration [@NosekRevolution2018].
Preregistration increases the credibility of empirical results by separating theory-driven planned analyses from unplanned exploratory analyses.
Preregistration seeks to reduce "researcher degrees of freedom" [@reference] in the execution of a planned study by limiting the number of decisions to be made after data has been collected, which could influence the results.
However, there are several shortcomings to preregistration.
One shortcoming is that written study plans are often multi-interpretable.
Even when several researchers set out to investigate the same hypothesis using the same data, their approaches vary considerably [@silberzahnManyAnalystsOne2018].
The current best practice to ensure that the preregistration are comprehensive and specific enough is to use preregistration templates [like @bowmanOSFPreregTemplate2020].
These templates seek to improve the specificity of study plans by imposing structure [@bakkerEnsuringQualitySpecificity2020].
However, templates cannot remove researchers degrees of freedoms completely,
because it is impossible to verbally describe every detail of an analysis for all but the simplest analysis.
A second limitation is that, just like primary research, preregistrations are susceptible to non-reproducibility.
To illustrate, a review of 210 preregistrations found that, even though 174 (67%) included a formal power analysis, only 34 (20%) of these could be reproduced [@bakkerRecommendationsPreregistrationsInternal2020].
Even when researchers have gone to great length in preregistering an analysis script,
they sometimes inexplicably fail to reproduce their own results.
For example, @steegenMeasuringCrowdAgain2014 realized after publication that part of their preregistered code results in different test statistic than they reported initially (see Footnote 7).
A final limitation is that written plans may turn out to be unfeasible once data are obtained and analyzed.
For example, a verbal description of a statistical model may be unidentified, e.g. if it includes reciprocal paths between variables, or more parameters than observed data.
Conversely, a model may be mis-specified in a major way; for example by omitting direct effects when the research question is about mediation, thus leading to a model with unacceptable fit.
Many researchers would only realize that such a model cannot be estimated once the data are obtained, thus necessitating a deviation from the preregistered plans.

@@

We believe that pairing PAC with the here presented workflow offers five advantages over classical preregistration.
First, in PAC, the preregistration is simply a first draft of the final manuscript,
thus sparing authors from writing (and editors and reviewers from evaluating) two separate documents.
Relatedly, writing a preregistration in form of a research article has the advantage that researchers already have great expertise in writing a scientific manuscript.
By contrast, filling out a preregistration template is a novel type of scientific writing for many.
Second, the PAC is a tool for study planning.
When all steps to be executed are documented clearly, the execution of the study will proceed more efficiently than when every step is planned in medias res.
Third, a PAC removes any ambiguity regarding the translation of verbal analysis plans into code.
Fourth, despite its rigor, PAC does accommodate data-dependent decisions if these can be formulated as code.
Researchers can, for example, formulate conditions (`if() {} else {}`) under which they prefer one analysis type over the other.
For example, if distributional assumptions are not met, the code may branch out to employ robust methods; or, an analysis may perform automated variable selection mechanisms before running the final model.
Fifth, a PAC is more comprehensive by design because its completeness can be empirically checked with simulated data.
Evaluating the intended analysis code on simulated data will help identify missing steps or ambiguous decisions.
PAC therefore minimizes researchers degrees of freedom more effectively than a standard preregistration [@bakkerEnsuringQualitySpecificity2020].
Sixth, deviations from the preregistration are clearly documented, because they are reflected in changes to the code which are managed and tracked with version control.
One could even make a commit labeled "Necessary deviations from the preregistration after data collection",
and reference it in the manuscript.

@@

When writing a PAC, researchers ought to strive to meet three goals:
First, the plan should be comprehensive, that is, researchers need to state the research question and describe the study design and analysis in detail.
Second, the plan should be effective, meaning that the researchers can reasonably expect it to answer the research question at hand.
Third, the plan should be efficient; the plan should only require the necessary amount of resources.

@@

Describing the preregistration in the form of a journal article requires authors to follow the same standards that also ensure that research articles are comprehensive enough to replicate the described study.
One particular advantage is the strong foundation in existing literature where the introduction and theoretical background provide the basis for the studies' design.
It can be fruitful to consult preregistration templates, because of there strong focus on formulating a comprehensive methods section [@ref provide specific good examples].

Another factor that ensures a comprehensive plan is writing the analysis based on simulated data.
Simulating data offers various benefits in the planning phase of a study.
It forces the researcher to think about and address data analytic problems on a conceptual level before they make the effort to acquire data [@axelrodAdvancingArtSimulation1997].
It also affords researchers with a simple technical check whether the analysis code runs without failure [@braiekTestingMachineLearning2020] (for example, does the code work with missing values, extreme outliers, etc.).
Writing incomplete code is technically impossible or at least researchers would notice if they e.g. forgot to aggregate the data etc.

@@

To write a PAC researchers should start by drafting a manuscript in form of an R Markdown.
When they have completed the method section, they should begin writing code.
We suggest writing three functions for each hypothesis, one for the planned analysis, one for the simulating the data and one for reporting the results.
At the heart of a PAC lies the planned analysis so researchers should first draft a function called `planned_analysis()` that captures the description in the method section.
This function should receive the data and derive relevant results from it.

@@

Because it is difficult to draft such function without interactively test it, researcher should than proceed to write a function for simulating data.

@@

When the researchers are satisfied with the function `planned_analysis()` they can think about how they would like to report the results of the analysis, via tables, plots and text.
The implementation of this reporting should be in the function `report_analysis()`.

```{r}
report_results <- function(results){
  # 1. create markdown tables from results
  # 2. conditionally interpret results e.g. if(p < .025)"Result is significant."
  # (optional) visualize results
  # 3. return results section formatted in markdown
}
```

These function again should accept the output of `planned_analysis()` as input.

The goal of this function is to describe the results to readers automatically.
This way, the preregistration specifies not only the computation  also how the computation's results are reported.
The output of this function should be formatted in Markdown.
A variety of packages automatically generate well formatted outputs of statistical reports or even entire tables of estimates or figures directly from R objects.
Packages like [`pander`](https://github.com/Rapporter/pander) [@pander], [`stargazer`](https://cran.r-project.org/web/packages/stargazer/vignettes/stargazer.pdf) [@stargazer], [`apaTables`](https://dstanley4.github.io/apaTables/articles/apaTables.html) [@apaTables] and [`papaja`](http://frederikaust.com/papaja_man/) [@papaja] help you to create dynamically generated professional looking results.
The package [`report`](https://github.com/easystats/report) [@report] is particularly noteworthy because it not only generates tables, but also a simple interpretation of the effects as actual prose (e.g it quantifies the size of an effect).

@@

To summarize researchers write three functions, `planned_analysis()`, `simulate_data()`, and `report_analysis()` and embed these into a manuscript that serves as preregistration.
After they gathered the actual data, they replace the simulated data, run `planned_analysis()` and write the discussion.

@@

To guide the reader through a preregistration as code (PAC), we will use an exemplary research question that was based on openly available data and merely chosen for simplicity of illustration:

> "Is there a mean difference in the personality trait 'Machiavellism' between self identified females and males?"

This example serves only didactic purposes and not to derive substantive claims.
Nevertheless, everything else is kept as close to a real research question as possible.

At the heart of a PAC lies the planned analysis.
We suggest writing a main analysis function that captures the full process of testing a research question.
This function receives a dataset as input and performs all statistical processing.
For the hypothetical research question, we could for example plan to conduct a simple Student's t-test [@studentProbableErrorMean1908] with Welch's correction [@welchGeneralizationStudentProblem1947] an the average of machiavellism items of the Short Dard Triad [SD3, @Jones2013] between the binary-coded gender groups.
As an example of a data-dependent decision, we included a rank transformation if the skewness was greater than 1, converting the t-test to a supposedly more robust Mann–Whitney–Wilcoxon test.[^caution-mww]

[^caution-mww]: The Mann–Whitney–Wilcoxon is in fact **not** more robust than the Welch corrected t-test for reasonably large sample sizes. The interested reader can use the provided code for the simulation to verify that the t-test provides unbiased effect sizes but the Mann–Whitney–Wilcoxon overestimates effect size with increasing sample size and skewness.

This function is one possible translation of the verbal description above:

@@

Importantly, natural language descriptions do not map unambiguously onto computer code implementations. 
The same textual descriptions may be implemented differently by different persons depending on their statistical and programming knowledge but also based on their own expectations and beliefs.
The present example illustrates this ambiguity, 
because the functions `wilcox.test` and the combinations of `rank` and `t.test` are both valid implementation of the Mann–Whitney–Wilcoxon test.
However, whereas the former assumes equal variance whereas the latter is robust even with unequal variances across groups [@zimmermanRankTransformationsPower1993].
Mentioning every such detail in a written preregistration plan is almost impossible.
Still, these details impact the results and provide unwanted researcher degrees-of-freedom.

```{r simulate_data}
```

Here, we decided to randomly draw non-normal data to test how well our procedure performed under non-optimal situations.
Specifically, we draw from a $\chi^2$-distribution, where in one group we add difference that conforms to a given standardized effect size (a Cohen's d of .2).
This simulated dataset fits perfectly into our planned analysis with a sample size of 400 and 10 items:

***

```{r}
set.seed(112358) # fix state of RNG
results <- 
  planned_analysis(
    simulate_data(
      n = 400,# Sample size
      df = 8, # Chisq with DF = 8 has skewness = 1
      d = .2, # Cohen's D
      i = 10  # Number of items (should not effect anything)
    )
  )
```

<!-- CJ Repetition from above!!!-->
Next, we add dynamic code blocks to our Markdown report that implement the planned reporting of the final results.
A variety of packages automatically generate well formatted outputs of statistical reports or even entire tables of estimates or figures directly from R objects.
Packages like [`pander`](https://github.com/Rapporter/pander) [@pander], [`stargazer`](https://cran.r-project.org/web/packages/stargazer/vignettes/stargazer.pdf) [@stargazer], [`apaTables`](https://dstanley4.github.io/apaTables/articles/apaTables.html) [@apaTables] and [`papaja`](http://frederikaust.com/papaja_man/) [@papaja] help you to create dynamically generated professional looking results.
The package [`report`](https://github.com/easystats/report) [@report] is particularly noteworthy because it not only generates tables, but also a simple interpretation of the effects as actual prose (e.g it quantifies the size of an effect).
Here we use the `report` package to create a preview of how our reporting will look like based on simulated data:<!--AB add citation to the report package-->

```{r, results='asis', eval=FALSE}
report::report(results$test)
```

This call to the `report()` function will yield the following automatically generated output:

```{r, results='asis', echo=FALSE}
report::report(results$test)
```
<!--AB it is difficult to discern the automatically generated text and out manuscript text even though there is the horizontal line. Not sure how to improve this.-->
***

@@

To determine a minimal sample size for this study we repeat this simulation across several combination of effect sizes and sample sizes.
One standard recomendations for choosing a sample size is to have an expected power of 80% based on the expected effect size.
@Jones2013 found in three studies gender differences of .24, .29, and .35 for their machiavilism subscale.
If we want to detect an effect of size `r choosen_d` with an `r choosen_power*100`% probability, our simulation shows that we require at least `r minn$n` participants.

>>>>

Scientists increasingly strive to make research data, materials, and analysis code openly available.
Sharing these types of digital research products has the potential to increase scientific efficiency by enabling researchers to learn from each other and reuse materials, and to increase scientific reliability by facilitating the review and replication of published research results.
To some extent these potential benefits are contingent on whether these digital research products are reproducible. 
Reproducibility can be defined as the ability of anyone to obtain identical results from the *same* data with the *same* computer code [see @Peikert2019 for details].
The credibility of empirical research hinges on its objectivity. Objectivity in this context means "in principle it can be tested and understood by anybody." [@popperLogicScientificDiscovery2002, p. 22][^popper].
Therefore, reproducibility has long been considered integral to empirical research.
Unfortunately, despite increasing commitment to open science practices and good will, many projects can not yet be reproduced by other research teams [@obels2020].
This is because making a research project with all code, data, and materials reproducible comes with a variety of challenges, and there is a dire need for reproducibility standards.
We now have the technology to make all digital products related to a research project available so that a research project can be reproduced with minimal effort.

This tutorial is aimed at researchers who intend to ensure the reproducibility of a research project and are willing to make relevant code, data, and materials available, whether publicly or on request.
We pursue two aims.

First, we introduce the reader to a workflow that maximizes the potential reproducibility of a research project.

This workflow builds on several technical solutions originating in the software engineering community. 
We discuss how these solution address four common threats to reproducibility [see @Peikert2019 for details] and form a coherent workflow.
We further demonstrate how the R package `repro` supports researchers in adopting this workflow.
Second, we discuss opportunities of reproducible workflows to improve not only how research outputs are created and recreated, but also how research itself is conducted.
One such opportunity is a strictly reproducible and unambiguous form of preregistration [@NosekRevolution2018] that builds upon a reproducible workflow, the so-called *preregistration as code* (PAC).
PAC involves preregistering the intended analysis code and the bulk of the manuscript as a dynamic document, including the Introduction, Methods, and Results sections.
The resulting dynamic document closely resembles the final manuscript, but uses simulated data to generate placeholder results (e.g. figures, tables and statistics).
Once the empirical data are analyzed, the Results are updated, and the Discussion is written.

@@

1.	**Multiple inconsistent versions of code, data, or both**; for example
  - the data set could have changed over time because outliers were removed at a later stage or an item was later recoded
  - the analysis code could have been modified during writing of a paper because a bug was removed at some point in time
It may then be unclear which version of code and data was used to produce some reported set of results.
2.	**Copy-and-paste errors**; for example, often, results are manually copied from a statistical computing language into a text processor; if a given analysis is re-run and results are manually updated in the text processor, this may inadvertently lead to inconsistencies between the reported result and the reproduced result;
3.	**Undocumented or ambiguous order of code execution**; for example, with multiple data and code files, it may be unclear which scripts should be executed in what order;  or, some of the computational steps are documented (e.g., final analysis), but other steps were conducted manually and not documented (e.g., running a scripts/commands by hand; copy-pasting results from one program to another)
4.	**Ambiguous dependencies**; for example, a given analysis may depend on a specific version of a specific software package, or rely on software that might not be available on a different computer or no longer exist at all; or, a different version of the same software may produce different results.

We have developed a workflow that achieves long-term and cross-platform computational reproducibility of scientific data analyses.
It leverages established tools and practices from software engineering and rests on four pillars that address the aforementioned causes of non-reproducibility: [@Peikert2019]:

@@

A version controlled dynamic document with dependency tracking may still rely on external software.
Troubleshooting issues specific to a particular programming language or dependent tool typically requires considerable expertise and pose a threat to reproducibility.
One solution to avoid problems caused by missing or inconsistent external software dependencies is to provide all dependent software packages and system libraries via **software management**.
One comprehensive approach to software management is containerization.
"packaging the key elements of the computational  environment  needed  to  run  the  desired  software [makes] the software much easier to use, and the results easier to reproduce" [@silverSoftwareSimplified2017, p. 174].

@@

1. Text (prose; e.g., a scientific paper or presentation)
2. Executable code (e.g., analyses)
3. Metadata (e.g. title, authors, document format)

R Markdown is a type of dynamic document well-suited to the RStudio user interface.
The text of an R Markdown is formatted by Markdown (see @xieMarkdownDefinitiveGuide2019 for technical details and @xieMarkdownCookbook2020 for practical guidance).
The code mostly consists of R-code (although other programming languages are supported, like Python, C++, Fortran, etc).
The following example serves to illustrate the Markdown syntax. 
It shows how to create a heading, a word in bold font, and a list of several items in Markdown:

@@

The metadata the `repro::automate()` function relies on can only express R packages as dependencies for the Dockerfile and only trivial dependencies (in the form of "file must exist") for the Makefile. 
The first limitation means that users must rely on software that is either already provided by the base Dockerimage "rocker/verse" or the R packages they list in the metadata.
Other software that users might need like other programming languages, not already installed LaTeX packages etc. have to be added manually.
We plan to add support for commonly used ways to install software beyond R packages via the metadata and `repro::automate()` e.g. for system libraries (via `apt` the Ubuntu package manager), LaTeX packages (via `tlmgr` the Tex Live package manager), python packages (via `pip` the python package manager).
The second limitation is about dependencies.
Make can represent complex dependencies like in this example: A depends on B, which in turn depends on C and D.
If B is missing in this example, Make would know how to recreate it from C and D.
These dependencies and especially how these dependencies should be resolved are difficult to represent in the metadata.
Users therefore have to either "flatten" the dependency structure, by simply stating that A depends on B, C and D, therefore leaving out important information or express the dependencies directly within the Makefile.

@@

Using Docker may require you to install software on an operating system that may not be familiar to you.
The images supplied by [@boettigerIntroductionRockerDocker2017], for example, are based on the Ubuntu operating system.
The most convenient way to install software on Ubuntu is through its package manager `apt`.

If the following snipped is added to `.repro/Dockerfile_manual` the Docker image will have Python installed.

@@

The file `.repro/Makefile_Rmds` contains the automatically generated targets from `repro::automate()` for the R Markdown files.
This file should not be altered manually.
If you are not satisfied with the automatically generated target simply provide an alternative target in the main `Makefile`.
Targets in the main `Makefile` take precedent.

@@

Defining the research questions and planning data analysis before observing the research outcomes is referred to as *preregistration* [@NosekRevolution2018].
Preregistration increases the credibility of empirical results by separating a-priori planned and theory-driven analyses from those conducted unplanned and post-hoc, that is, after having seen the data.
Preregistration seeks to reduce "researcher degrees of freedom" [@reference] in the execution of a planned study by limiting the number of decisions to be made after data has been collected, which could influence the results.
Preregistrations, however, have several shortcomings.
One shortcoming is that written study plans are often multi-interpretable and often do not achieve the level of unambiguity that are required to effectively zero the researcher degrees of freedom.
Even when several researchers describe their analysis with the same terms, use the same data, and investigate the same hypothesis, the results vary considerably [@silberzahnManyAnalystsOne2018].
The current best practice to ensure that the preregistration are comprehensive and specific enough is to use preregistration templates [like @bowmanOSFPreregTemplate2020].
These templates seek to improve the specificity of study plans by imposing structure [@bakkerEnsuringQualitySpecificity2020].
However, templates cannot remove researchers degrees of freedoms completely, because it is impossible to verbally describe every detail of an analysis for all but the most straightforward analysis.
A second limitation is that, just like primary research, preregistrations are susceptible to non-reproducibility.
To illustrate, a review of 210 preregistrations found that, even though 174 (67%) included a formal power analysis, only 34 (20%) of these could be reproduced [@bakkerRecommendationsPreregistrationsInternal2020].
Even when researchers have gone to great length in preregistering an analysis script,
they sometimes inexplicably fail to reproduce their own results.
For example, @steegenMeasuringCrowdAgain2014 realized after publication that part of their preregistered code results in different test statistic than they reported initially (see Footnote 7).
A final limitation is that written plans may turn out to be unfeasible once data are obtained and analyzed.
For example, a verbal description of a statistical model may be unidentified, e.g. if it includes reciprocal paths between variables, or more parameters than observed data.
Conversely, a model may be mis-specified in a major way; for example by omitting direct effects when the research question is about mediation, thus leading to a model with unacceptable fit.
Many researchers would only realize that such a model cannot be estimated once the data are obtained, thus necessitating a deviation from the preregistered plans.

@@

We believe that pairing PAC with the here presented workflow offers five advantages over classical preregistration.
First, in PAC, is merely an intermediate stage of the final manuscript, thus sparing authors from writing, and editors and reviewers from evaluating two separate documents.
Relatedly, writing a preregistration in form of a research article has the advantage that researchers already have great expertise in writing a scientific manuscript.
By contrast, filling out a preregistration template is a novel type of scientific writing for many.
Second, the PAC is a tool for study planning.
When all steps to be executed are documented clearly, the execution of the study will proceed more efficiently than when every step is planned in medias res.
Third, a PAC removes any ambiguity regarding the translation of verbal analysis plans into code.
A PAC is more comprehensive by design because its completeness can be empirically checked with simulated data.
Evaluating the intended analysis code on simulated data will help identify missing steps or ambiguous decisions.
PAC therefore minimizes researchers degrees of freedom more effectively than a standard preregistration [@bakkerEnsuringQualitySpecificity2020].
Fourth, despite its rigor, PAC does accommodate data-dependent decisions if these can be formulated as code.
Researchers can, for example, formulate conditions (`if() {} else {}`) under which they prefer one analysis type over the other.
For example, if distributional assumptions are not met, the code may branch out to employ robust methods; or, an analysis may perform automated variable selection mechanisms before running the final model.
Another example of data dependent decisions are more explorative analyses, i.e. explorative factor analysis or machine learning.
With PAC researchers can precisely calibrate how much they want to explore.
When communicating a finding as explorative, it may still make a difference if tens or hundreds of variables were considered.
Fifth, deviations from the preregistration are clearly documented, because they are reflected in changes to the code which are managed and tracked with version control.
One could even make a commit labeled "Necessary deviations from the preregistration after data collection",
and reference it in the manuscript.

@@

When planning a study and hence when writing a PAC, researchers ought to strive to meet three goals:
First, the plan should be comprehensive, that is, researchers need to state the research question and describe the study design and analysis in detail.
Second, the plan should be effective, meaning that the researchers can reasonably expect it to answer the research question at hand.
Third, the plan should be efficient; the plan should only require the necessary amount of resources.

@@

Describing the preregistration in form of a journal article requires authors to follow the same standards that also ensure that research articles are comprehensive enough to replicate the described study.
One particular advantage is the foundation in existing literature where the introduction and theoretical background provide the basis for the studies' design.
Nevertheless, it can be fruitful to consult preregistration templates because of their strong focus on formulating a comprehensive methods section.

Another factor that ensures a comprehensive plan is writing the analysis based on simulated data.
Simulating data offers various benefits in the planning phase of a study.
It forces the researcher to think about and address data analytic problems on a conceptual level before they make the effort to acquire data [@axelrodAdvancingArtSimulation1997].
It also affords researchers with a simple technical check whether the analysis code runs without failure [@braiekTestingMachineLearning2020] (for example, does the code work with missing values, extreme outliers, etc.).
Writing incomplete code is technically impossible, or at least researchers would notice if something is omitted (e.g. if they forgot to aggregate the data).

@@

To write a PAC, researchers should start by drafting a manuscript in the form of an R Markdown.
When they have completed the method section, they should begin writing code.
We suggest writing three functions for each planned hypothesis, one for the planned analysis, one for simulating the data and one for reporting the results.
At the heart of a PAC lies the planned analysis, so researchers should first draft a function called `planned_analysis()` that fits the description in their method section.
This function should receive the data and derive relevant results from it.

@@

@@

When the researchers are satisfied with the function `planned_analysis()`, they can think about how they would like to report the analysis results via tables, plots, and text.
The implementation of this reporting should be in the function `report_analysis()`.

```{r, eval=FALSE}
report_analysis <- function(results){
  # 1. create markdown tables from results
  # 2. conditionally interpret results e.g. if(p < .025)"Result is significant."
  # (optional) visualize results
  # 3. return results section formatted in markdown
}
```

This function should again accept the output of `planned_analysis()` as input.
The output of this function should be formatted in Markdown.
The idea is to automatically generate the full results section from the analysis.
This way, the preregistration specifies not only the computation but also how the computation's results are reported.
Various packages automatically generate well-formatted Markdown outputs of statistical reports or even entire tables of estimates or figures directly from R objects to help with this objective.
Packages like [`pander`](https://github.com/Rapporter/pander) [@pander], [`stargazer`](https://cran.r-project.org/web/packages/stargazer/vignettes/stargazer.pdf) [@stargazer], [`apaTables`](https://dstanley4.github.io/apaTables/articles/apaTables.html) [@apaTables] and [`papaja`](http://frederikaust.com/papaja_man/) [@papaja] help you to create dynamically generated professional looking results.
The package [`report`](https://github.com/easystats/report) [@report] is particularly noteworthy because it generates not only tables but also a straightforward interpretation of the effects as actual prose (e.g. it verbally quantifies the size of an effect).

@@

To summarize, researchers write three functions, `planned_analysis()`, `simulate_data()`, and `report_analysis()` and embed these into a manuscript that serves as a preregistration.
After they gathered the actual data, they replace the simulated data, render the dynamic manuscript (therefore run `planned_analysis()` on the actual data) and write the discussion.

@@

We argued that PAC has several advantages over classical preregistration and have outlined how to implement it.
To illustrate how PAC works in practice and help researchers implementing PAC themselves, we have conducted a PAC that serves as a guiding example.
We will use an exemplary research question that was based on openly available data and merely chosen for simplicity of illustration:

> "Is there a mean difference in the personality trait 'Machiavellism' between self-identified females and males?"

Again, we propose a preregistration format that closely resembles a classical journal article, but uses simulated data and dynamic document generation to create a document that starts out as a preregistration and eventually becomes the final report.
The complete preregistration source is available in the [online supplementary material](https://github.com/aaronpeikert/repro-tutorial/blob/main/preregistration.Rmd).
In this section, we show code excerpts of this preregistration (formatted in monospace) and explain the rationale behind them.

```{r, echo=FALSE}
show <- function(x, wrap = TRUE, cat = TRUE, fence = TRUE){
  stopifnot(length(x) == 1L)
  if(wrap)x <- str_c(str_wrap(str_split(x, "\\n\\n")[[1]]), collapse = "\n\n")
  if(!isFALSE(fence)){
    if(isTRUE(fence))fence <- "markdown"
    fence <- str_c("````", fence)
    x <-
      str_c(fence,
            x,
            "````",
            sep = "\n",
            collapse = "\n")
  }
  if(cat){
    cat(x)
    return(invisible(x))
  } else {
    return(x)
  }
}
show_prereg <-
  function(section,
           file = "preregistration.Rmd",
           wrap = TRUE,
           cat = TRUE,
           fence = TRUE) {
    prereg <- readr::read_lines(here::here(file))
    start <- stringr::str_c("<!--begin-", section, "-->")
    end <- stringr::str_c("<!--end-", section, "-->")
    start_pos <- which(prereg == start)
    end_pos <- which(prereg == end)
    if (length(start_pos) != 1L || length(end_pos) != 1L)
      stop("No section '", section, "' found.")
    excerpt <-
      str_c(prereg[(start_pos + 1):(end_pos - 1)], collapse = "\n")
    show(excerpt, wrap = wrap, cat = cat, fence = fence)
  }
```

As usual, researchers state why they are interested in their research question in the section "Introduction" and provide the necessary background information and literature to understand the context and purpose of the research question.
This example is drastically shortened for illustration purposes:

```{r, echo=FALSE, results='asis'}
show_prereg("theory")
```

After researchers have provided the research question, they typically proceed to explain how they want to study it.
For simplicity, we will use already published data that we have not yet analyzed:

```{r, echo=FALSE, results='asis'}
show_prereg("method")
```

We choose the following statistical procedure because many researchers are familiar with it[^caution-mww]:

```{r, echo=FALSE, results='asis'}
show_prereg("ttest")
```

[^caution-mww]: The Mann–Whitney–Wilcoxon is in fact **not** more robust than the Welch corrected t-test for reasonably large sample sizes. The interested reader can use the [provided code for the simulation](https://github.com/aaronpeikert/repro-tutorial/blob/main/simulation.R) to verify that the t-test provides unbiased effect sizes but the Mann–Whitney–Wilcoxon overestimates effect size with increasing sample size and skewness.

We then translated this methods section into the following `planned_analysis()` function:

@@

This function illustrates two advantages of PAC.
First, a PAC can easily include data-dependent decisions creating different analysis branches under different conditions.
Second, it highlights how difficult it is to describe a statistical analysis precisely.
The same verbal descriptions may be implemented differently by different persons depending on their statistical and programming knowledge and their expectations and beliefs.
One example would be using the function `wilcox.test` instead of the combinations of the functions `rank` and `t.test`.
Either of them are valid implementations of the Mann–Whitney–Wilcoxon test, but the first assumes equal variance.
In contrast, the second applies Welch's correction by default and hence is robust even with unequal variances across groups [@zimmermanRankTransformationsPower1993].
Mentioning every such minute implementation detail is almost impossible and would result in overly verbose preregistrations.
Still, these details can make a difference and provide unwanted researchers' degrees of freedom.

The function `planned_analysis()` together with the function `simulate_data()` (not shown here) can be used to justify the planned sample size.
To that end, `simulate_data()` is repeatedly called with increased sample size and the proportion of significant results (power) is recorded.
The results for such a Monte Carlo simulation for this example are visualized in figure \@ref(fig:power).
The code for this power analysis can be found in the [online supplementary material](https://github.com/aaronpeikert/repro-tutorial/blob/main/simulation.R).

Monte Carlo simulations are, of course, not only applicable for this analysis method and also allow researchers to investigate other relevant properties of their analysis method beyond power [@brandmaier2018precision; @harrisonIntroductionMonteCarlo2010; @raychaudhuriIntroductionMonteCarlo2008].

@@

```{r, include=FALSE}
fig_caption <- str_c("Results of simulation for the power analysis. The cross indicates the sample size that archives ", choosen_power*100, "% assuming a Cohen's D of ", choosen_d, ".")
```

```{r power, echo=FALSE, fig.cap=fig_caption}
res %>%
  ggplot(aes(n, power, color = d, group = d)) +
  geom_line() +
  geom_point(
    data = minn,
    size = 3,
    color = "black",
    shape = 3
  ) +
  theme_minimal() +
  scale_color_viridis_c() +
  theme(legend.title = element_text()) +
  labs(y = "Power\n(Proportion of significant results)",
       x = "Sample size",
       color = "Cohen's D") +
  NULL
```

```{r, echo=FALSE, results='asis'}
show_prereg("power-code", wrap = FALSE)
```

```{r, echo=FALSE, results='asis'}
show_prereg("power")
```
